---
title: "HuanHusted_TSA_Competition"
author: "Kelsey Husted & Yu Huan"
date: "2023-04-28"
output: pdf_document
---

## Time Series Competition

```{r setup,warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

##Load packages

```{r load packages, warning=FALSE, message=FALSE}
library(readxl)
library(dplyr)
library(lubridate)
library(openxlsx)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(zoo)
library(kableExtra)
```

## Data wrangling and processing from 2005-2009

Hourly data was transformed into daily data with aggregate functions and pipes (i.e., tidyverse).

```{r Wrangle & Process Data 2005-2009, message=FALSE, warning=FALSE}
#Import data
df <- read_xlsx("./Competition/Data/load.xlsx")

df_T <- read_xlsx("./Competition/Data/temperature.xlsx")

df_H <- read_xlsx("./Competition/Data/relative_humidity.xlsx")

#Wrangle data from hourly to daily
#Wrangling date column 2005 to 2009
df_daily <- df %>%
  mutate( Date = ymd(date)) %>%
  filter(Date < '2010-01-01')

#removing no numeric columns so rowMeans() function will work
df_dailyV2 <- df %>%
  mutate( Date = ymd(date)) %>%
  filter(Date < '2010-01-01') %>%
  select(3:26)

#Creating daily data
df_processed <- df_dailyV2 %>%  
  mutate(rowMeans(df_dailyV2)) %>%
  rename(Daily_data = "rowMeans(df_dailyV2)") %>%
  select(25)

#Combining date and daily data
date <- df_daily[,2]
df_processed <- cbind(date, df_processed)

nobs = nrow(df_daily)           
```

## Data wrangling and processing from 2005-2010

The data needs to be formatted to include 2010 as well since the objective is to forecast for 2011.
Instead of making two separate datasets, I should use the window() function for future reference.

```{r Wrangle & Process Data 2005-2010, message=FALSE, warning=FALSE}

#Wrangle data from hourly to daily
#Wrangling date column 2005 to 2010
#removing no numeric columns so rowMeans() function will work
df_daily2010 <- df %>%
  mutate( Date = ymd(date)) %>%
  select(3:26)

#Creating daily data
df_processed2010 <- df_daily2010 %>%  
  mutate(rowMeans(df_daily2010)) %>%
  rename(Daily_data = "rowMeans(df_daily2010)") %>%
  select(25)

#Combining data and daily data
date <- df[,2]
df_processed2010 <- cbind(date, df_processed2010)

nobs2010 = nrow(df_processed2010)        
```

## Time series object transformation

```{r ts, message=FALSE, warning=FALSE}
#ts transformation 2005 to 2009
ts_daily <- msts(df_processed$Daily_data,
                 seasonal.periods=c(7,365.25),
                 start=c(2005, 01, 01))

#ts transformation 2005 to 2010
ts_daily2010 <- msts(df_processed2010$Daily_data,
                 seasonal.periods=c(7,365.25),
                 start=c(2005, 01, 01))

#ts for accuracy test visalization

ts_daily2010_test <- msts(df_processed2010$Daily_data[1:1885],
                 seasonal.periods=c(7,365.25),
                 start=c(2005, 01, 01))

```

## Time series decomposition and plot

```{r Decompose, message=FALSE, warning=FALSE}

#Decompose time series 
ts_decompose <- ts_daily %>%
  mstl()
plot(ts_decompose)

```
## Model 1 STL + ETS: Forecast 2010

```{r Model 1, message=FALSE, warning=FALSE}
#Fit and forecast STL model 
ETS_model <-  stlf(ts_daily,h=59)

#Plot foresting
autoplot(ETS_model) + ylab("Forecast")

#Plot model + observed data
autoplot(ts_daily2010_test) +
  autolayer(ETS_model, series="STL Model",PI=FALSE) +
  ylab("Daily Demand")


#Plot the residuals

checkresiduals(ETS_model)


#Check accuracy of model
n_for <- 59
observed <- df_processed2010[1827:1885, "Daily_data"]
ETS_scores <- accuracy(ETS_model$mean,observed)
print(ETS_scores)




```

## Model 1 STL + ETS: Forecast 2011

```{r Model 1 2011, message=FALSE, warning=FALSE}
#Fit and forecast STL model January 1st to February 28th 2011
ETS_model2011 <-  stlf(ts_daily2010,h=59)

#Plot foresting
autoplot(ETS_model2011) + ylab("Forecast")

#Plot model + observed data
autoplot(ts_daily2010) +
  autolayer(ETS_model2011, series="STL Model",PI=FALSE) +
  ylab("Daily Demand")


```


## Model 2 TBATS: Forecast 2010

# The model looks like a really bad fit visually and will not be used to forecast for 2011.

```{r Model 2, message=FALSE, warning=FALSE}
#Fit and forecast TBATS model 
TBATS_model <-  tbats(ts_daily)

#forecast
TBATS_for <- forecast(TBATS_model,h=59)

#Plot foresting
autoplot(TBATS_for) + ylab("Forecast")

#Plot the residuals

checkresiduals(TBATS_model)

#Plot model + observed data
autoplot(ts_daily2010_test) +
  autolayer(TBATS_for, series="TBATS Model",PI=FALSE) +
  ylab("Daily Demand")


#Check accuracy of model
n_for <- 59
observed <- df_processed2010[1827:1885, "Daily_data"]
TBATS_scores <- accuracy(TBATS_for$mean,observed)
print(TBATS_scores)


```

## Model 2 TBATS: Forecast 2011


```{r}
#Fit and forecast TBATS model 
TBATS_model2011 <-  tbats(ts_daily2010)

#forecast
TBATS_for2011 <- forecast(TBATS_model2011,h=59)

#Plot foresting
autoplot(TBATS_for2011) + ylab("Forecast")

#Plot the residuals

checkresiduals(TBATS_model2011)

#Plot model + observed data
autoplot(ts_daily2010) +
  autolayer(TBATS_for2011, series="TBATS Model",PI=FALSE) +
  ylab("Daily Demand")


```



## Model 3 ARIMA + FOURIER terms: Forecast 2010

```{r Model 3, message=FALSE, warning=FALSE}
#Fit and forecast TBATS model 
ARIMA_Four_model <- auto.arima(ts_daily, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_daily, 
                                          K=c(2,12))
                             )

#Forecast
ARIMA_Four_for <- forecast(ARIMA_Four_model,
                           xreg=fourier(ts_daily,
                                        K=c(2,12),
                                        h=59),
                           h=59
                           ) 

#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Daily Demand")

#Plot model + observed data
autoplot(ts_daily2010_test) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Daily Demand")

# Plot the residuals

checkresiduals(ARIMA_Four_model)

#Check accuracy of model
n_for <- 59
observed <- df_processed2010[1827:1885, "Daily_data"]
ARIMA_Four_scores <- accuracy(ARIMA_Four_for$mean,observed)
print(ARIMA_Four_scores)



```

## Model 3 ARIMA + FOURIER terms: Forecast 2011

```{r Model 3 2011, message=FALSE, warning=FALSE}
#Fit and forecast TBATS model 
ARIMA_Four_model2011 <- auto.arima(ts_daily2010, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_daily2010, 
                                          K=c(2,12))
                             )

#Forecast
ARIMA_Four_for2011 <- forecast(ARIMA_Four_model2011,
                           xreg=fourier(ts_daily2010,
                                        K=c(2,12),
                                        h=59),
                           h=59
                           ) 

#Plot foresting results
autoplot(ARIMA_Four_for2011) + ylab("Daily Demand")

#Plot model + observed data
autoplot(ts_daily2010) +
  autolayer(ARIMA_Four_for2011, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Daily Demand")




#print(ARIMA_Four_for2011$mean)
```


## Model 4 Neural Network Time Series: Forecasts 2010

```{r Model 4 2010, message=FALSE, warning=FALSE}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_model <- nnetar(ts_daily,decay=0.5, maxit=150, p=1,P=0,xreg=fourier(ts_daily, K=c(2,12)))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_model, h=59,xreg=fourier(ts_daily, 
                                          K=c(2,12),h=59))

#Plot foresting results
autoplot(NN_for) +
  ylab("Daily Demand") 

#Plot model + observed data
autoplot(ts_daily2010_test) +
  autolayer(NN_for, series="Daily Demand",PI=FALSE)+
  ylab("Daily Demand") 

checkresiduals(NN_model)

#Checking error variables to decide which model fits the data the best
n_for <- 59
observed <- df_processed2010[1827:1885, "Daily_data"]
NN_scores1 <- accuracy(NN_for$mean,observed)
print(NN_scores1)



```

## Model 4 Neural Network Time Series: Forecasts 2011

 > Note: Based on the error variables calculated with the accuracy() function, the Neural Network model seems to fit the data the best.

```{r Model 4 2011, message=FALSE, warning=FALSE}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_model2010 <- nnetar(ts_daily2010,lambda = 0.5,p=1,P=0,xreg=fourier(ts_daily2010, K=c(2,12)))

#NN_for <- forecast(NN_fit, h=365) 
NN_for2010 <- forecast(NN_model2010,PI=TRUE, h=59,xreg=fourier(ts_daily, 
                                          K=c(2,12),h=59))

#Plot foresting results
autoplot(NN_for2010) +
  ylab("Daily Demand") 

#Plot model + observed data
autoplot(ts_daily2010) +
  autolayer(NN_for2010, series="Daily_data",PI=FALSE)+
  ylab("Daily Demand") 



print(NN_for2010$mean)

print(ARIMA_Four_for2011$mean)





```


## Compare performance matrix

```{r}

#create data frame
seas_scores <- as.data.frame(rbind(ETS_scores, TBATS_scores, ARIMA_Four_scores, NN_scores1))
row.names(seas_scores) <- c("STL-ETS","TBATS", "ARIMA_FOUR","NEU-NETWORK")

#choose model with lowest RMSE
best_model_index <- which.min(seas_scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(seas_scores[best_model_index,])) 


kbl(seas_scores, 
      caption = "Forecast Accuracy for Seasonal Data",
      digits = array(5,ncol(seas_scores))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores[,"RMSE"]))


```



